Design and Implementation of Stable High Power Optical Lattices for Quantum Gas Microscopy

### 2017-11-12 LatticeOptics/latticeOpticsV2.tex
added figure
V4 of paper

\begin{document}

@@

\date{\today}
\begin{abstract}
We describe the design of a stable high power 1064 nm laser for use in optical lattice experiments. The system is based on a high quality mephisto NPRO seeding an array of four heavily modified fiber amplifiers. The intensity of every beam is stabilized with a low noise, nonlinear feedback system, and the waist can be smoothly controlled. 
\end{abstract}
\maketitle
\section{Introduction}
Ultracold atoms in optical lattices have become a powerful platform for experiments ranging from investigations of strongly correlated systems, quantum state engineering, quantum computing architectures and precision metrology, among others.
Further, the recent advent of quantum gas microscopy has enabled studies that exhibit unprecedented control over few atom systems that function at extremely low energy scales.
However, this rapid progress places ever more stringent technical constraints on the lasers used to trap and manipulate the atomic systems.
For example, in experiments at low temperatures, heating due to the intensity noise of a lattice cannot be allowed to approach the thermalization timescales lest the system fail to thermalize at the required temperature.
Similarly, experiments that rely on a small number of atoms held by multiple traps places stringent requirements on the relative stability of the traps.
We demonstrate a high power lattice system that exhibits levels of intensity noise below -120 dBc between 1 kHz and 3 MHz, and is positionally stable at the XXX site level. 
This system has been successfully used in a Fermi gas microscope to investigate the charge and spin sectors of the Hubbard model with unprecedented control.

@@

\section{Nufern Fiber Amplifiers}
Quantum gas microscopy of Li requires single site trap frequencies of approximately 1 MHz \cite{single site paper}.
Such trap frequencies may only be attained using tightly focused high power beams, which requires that the seed power be amplified to approximately 30 W per lattice axis, which is done using Nufern Fiber amplifiers (PART NUMBER XXX).
These amplifiers can amplify 150 mW to up to approximately 45 W (depending on pump diode current) by using a length of fiber doped with XXX (erbium or ytterbium) and pumped with xxx nm light from fiber coupled diode sources to provide a gain medium for 1064 nm light. 
The amplification is performed using a two stage amplifier design, where the stages are contolled using separate power supplies and electronics.
The high powers supplied by the fiber amplifiers raise two engineering difficulties. First, the fiber amplification process is inherently noisy, due to acoustic coupling of the fiber to the environment, electrical design of the amplifier, and spontaneous Brillouin scattering (SBS).

@@

\subsection{Residual Intensity Noise}
RIN:
\begin{itemize}
\item what is it and how do you measure it
\item easy to measure, enters into parametric heating effects
\end{itemize}

@@

\subsection{Fiber Amplifier Modifications}
FIGURE - SBS DATA
\begin{itemize}
	\item Mephisto+Fiber amplifiers
\item fiber amplifiers work like so so
\item schematic what it looks like and does
\item Modifications
	\begin{itemize}
		\item electronics control board,
		\item PSU's
		\item filters
		\item switching spikes removed
		\item fan noise stopped
	\end{itemize}
\item Two stage amplifier
	\begin{itemize}
		\item second stage has this special fiber 
		\item large mode area fibers
		\item leakage into cladding modes 
		\item sbs (cut off)
	\end{itemize}
\item SBS data
\end{itemize}
\section{Beam Monitoring}
A major recent direction of ultracold atoms experiments is the ability to use very few, individually-addressible atoms\cite{AFM paper, quantum walks few body} that are simultaneously held by several lattice axes in addition to auxilliary traps.
Unfortunately, such experiments require superb stability, where the underlying harmonic confinement does not shift by more than approximately 2-3 micron from experiment to experiment.

>>>>

\graphicspath{{./figures/}}
\begin{document}

@@

\date{\today}
\begin{abstract}
We describe the design of a stable high power 1064 nm laser for use in optical lattice experiments. The system is based on a high quality mephisto NPRO seeding an array of four heavily modified fiber amplifiers. The intensity of every beam is stabilized with a low noise, nonlinear feedback system, and the waist can be smoothly controlled. 
\end{abstract}
\maketitle
\section{Introduction}
Ultracold atoms in optical lattices have become a powerful platform for experiments ranging from investigations of strongly correlated systems, quantum state engineering, quantum computing architectures and precision metrology, among others.
Further, the recent advent of quantum gas microscopy has enabled studies that exhibit unprecedented control over few atom systems that function at extremely low energy scales.
However, this rapid progress places ever more stringent technical constraints on the lasers used to trap and manipulate the atomic systems.
For example, in experiments at low temperatures, heating due to the intensity noise of a lattice cannot be allowed to approach the thermalization timescales lest the system fail to thermalize at the required temperature.
Similarly, experiments that rely on a small number of atoms held by multiple traps places stringent requirements on the relative stability of the traps.
We demonstrate a high power lattice system that exhibits levels of intensity noise below -120 dBc between 1 kHz and 3 MHz, and is positionally stable at the XXX site level. 
This system has been successfully used in a Fermi gas microscope to investigate the charge and spin sectors of the Hubbard model with unprecedented control. 
\begin{figure}
    \includegraphics{figures/figure1/figure1.pdf}
    \caption{High level overview showing the major components of the lattice laser feedback system}
\end{figure}

@@

\section{Nufern Fiber Amplifiers}
Quantum gas microscopy of Li requires single site trap frequencies of approximately 1 MHz \cite{Parsons2015}.
Such trap frequencies may only be attained using tightly focused high power beams, which requires that the seed power be amplified to approximately 30 W per lattice axis, which is done using Nufern Fiber amplifiers (PART NUMBER XXX).
These amplifiers can amplify 150 mW to up to approximately 45 W (depending on pump diode current) by using a length of fiber doped with XXX (erbium or ytterbium) and pumped with xxx nm light from fiber coupled diode sources to provide a gain medium for 1064 nm light. 
The amplification is performed using a two stage amplifier design, where the stages are contolled using separate power supplies and electronics.
The high powers supplied by the fiber amplifiers raise two engineering difficulties. First, the fiber amplification process is inherently noisy, due to acoustic coupling of the fiber to the environment, electrical design of the amplifier, and spontaneous Brillouin scattering (SBS).

@@

\subsection{Residual Intensity Noise}
RIN:
\begin{itemize}
\item what is it and how do you measure it
\item easy to measure, enters into parametric heating effects\cite{Savard1997}
\end{itemize}

@@

\subsection{Fiber Amplifier Modifications}
FIGURE - SBS DATA
\begin{itemize}
	\item Mephisto+Fiber amplifiers
\item fiber amplifiers work like so so
\item schematic what it looks like and does
\item Modifications
	\begin{itemize}
		\item electronics control board,
		\item PSU's
		\item filters
		\item switching spikes removed
		\item fan noise stopped
	\end{itemize}
\item Two stage amplifier
	\begin{itemize}
		\item second stage has this special fiber 
		\item large mode area fibers
		\item leakage into cladding modes 
		\item sbs (cut off)
	\end{itemize}
\item SBS data
\end{itemize}
\section{Beam Monitoring}
A major recent direction of ultracold atoms experiments is the ability to use very few, individually-addressible atoms\cite{Mazurenko2016, Farhi} that are simultaneously held by several lattice axes in addition to auxilliary traps.
Unfortunately, such experiments require superb stability, where the underlying harmonic confinement does not shift by more than approximately 2-3 micron from experiment to experiment.
